---
title: "lobster_exploration"
author: "Marie Rivers"
date: "8/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
```
# Task 2: SBC Lobsters
```{r}
lobster_abundance <- read_csv(here("data", "Lobster_Abundance_All_Years_20210412.csv"), na = c("-99999", "")) 

lobster_abundance <- lobster_abundance %>% 
  clean_names()
```
Use `dplyr::uncount()` to convert the data from the *frequency* of a thing to each *case* of the thing
```{r}
lobster_abundance_case <- lobster_abundance %>% 
  uncount(count)
# this fiction removed the "count" column. If an observation recorded a count of 4, then that observation was converted into 4 rows
```

Find counts and mean sizes by site and year
```{r}
lobster <- lobster_abundance_case %>% 
  group_by(site, year) %>% 
  summarise(count = n(),
            mean_carapace_length = mean(size_mm, na.rm = TRUE))
lobster
```
Create a ggplot graph of the number of total lobsters observed (y-axis) by year (x-axis) in the study, grouped (either aesthetically or by faceting) by site
```{r}
ggplot(lobster, aes(x = year, y = count)) +
  geom_point(aes(color = site)) +
  facet_wrap(vars(site))
  labs (x = "year",
         y = "# of lobsters observed")
```
## Find the proportion of legal lobsters at each site for 2020  
The legal lobster size (carapce length) in California is 79.76 mm.
```{r}
lobster_abundance_case <- lobster_abundance_case %>% 
  mutate(legal = case_when(
    size_mm >= 79.76 ~ "yes",
    size_mm <79.76 ~ "no"))
  
legal_lobsters <- lobster_abundance_case %>% 
  filter(size_mm >= 79.76) %>% 
  filter(year == "2020")

# find the count of lobsters observed at each site (only using site as the grouping factor) that are above and below the legal limit.
legal_summary <- lobster_abundance_case %>% 
  group_by(site, legal) %>% 
  filter(legal != "na") %>% 
  filter(year == "2020") %>% 
  summarise(count_legal = n())
legal_summary
```
Create a stacked column graph that shows the proportion of legal and non-legal lobsters at each site
```{r}
ggplot(legal_summary, aes(x = site, y = count_legal)) +
  geom_col(aes(fill = legal), position = "fill")
```
## Task 3: Random lobster wrangling
filter() practice
```{r}
ex_a <- lobster_abundance %>% 
  filter(site %in% c("IVEE", "CARP", "NAPL"))

ex_b <- lobster_abundance %>% 
  filter(month == "8")

ex_c <- lobster_abundance %>% 
  filter(site == "AQUE" | size_mm > 70)

ex_d <- lobster_abundance %>% 
  filter(site != "NAPL")
```
group_by() %>% summarize() practice
```{r}
ex_e <- lobster_abundance %>% 
  group_by(site) %>% 
  summarise(mean_carapace = mean(size_mm, na.rm = TRUE),
            sd_carapace = sd(size_mm, na.rm = TRUE))
ex_e

ex_f <- lobster_abundance %>% 
  group_by(site, month) %>% 
  summarize(max_carapace = max(size_mm, na.rm = TRUE))
ex_f
```
mutate() practice
```{r}
ex_g <- lobster_abundance %>% 
  mutate(size_cm = size_mm / 100)
ex_g

ex_h <- lobster_abundance %>% 
  mutate(site = tolower(site))
ex_h

ex_i <- lobster_abundance %>% 
  mutate(area = as.character(area))
ex_i
is.numeric(ex_i$area)

```
case_when() practice
```{r}
ex_j <- lobster_abundance %>% 
  mutate(size_bin = case_when(
    size_mm <= 70 ~ "small",
    size_mm > 70 ~ "large"))
ex_j

ex_k <- lobster_abundance %>% 
  mutate(designation = case_when(
    site == "IVEE" ~ "MPA",
    site == "NAPL" ~ "MPA",
    TRUE ~ "not MPA"))
ex_k
```
